---
title: "South East Michigan Covid-19 Report"
author: "Roger Pique-Regi"
date: "3/29/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(usmap)
```

Data downloaded from <https://github.com/nytimes/covid-19-data> assembled by The New York Times. Rmd file to generate this is in <https://github.com/piquelab/progress-covid-19>. The plots are similar to those reported earlier by <https://www.ft.com/coronavirus-latest>.

```{r echo=FALSE, message=FALSE, warning=FALSE}
dc = read_csv("https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv")
cat("Last Update:",as.character.Date(select(dc[nrow(dc),],date))[4],"\n")
```

Counties aggregated: "Wayne","Oakland","Macomb","Livingston","St. Clair","Lapeer","Genesee","Washtenaw","Monroe","Lenawee"

```{r echo=FALSE}
dd <- dc %>% filter(state=="Michigan",county %in% c("Wayne","Oakland","Macomb","Livingston","St. Clair","Lapeer","Genesee","Washtenaw","Monroe","Lenawee")) %>% group_by(fips) %>% summarise_all(last) %>% left_join(countypop,by="fips") %>% mutate(drate=deaths/cases*100,dpoprate=deaths/pop_2015*1E6,cpoprate=cases/pop_2015*1E6)
plot_usmap(data=dd,values="drate",include="MI") + labs(fill = "%Deaths/Cases") 
#plot_usmap(data=dd,values="dpoprate",include=dd$fips) + labs(fill = "Deaths/Million(pop)") 
#plot_usmap(data=dd,values="cpoprate",include=dd$fips) + labs(fill = "Cases/Million(pop)") 
```

## South East Michigan Deaths
```{r echo=FALSE}
dd <- dc %>% filter(state=="Michigan",county %in% c("Wayne","Oakland","Macomb","Livingston","St. Clair","Lapeer","Genesee","Washtenaw","Monroe","Lenawee")) %>% group_by(date) %>% summarize(deaths=sum(deaths),cases=sum(cases)) %>% filter(deaths>=10) %>% mutate(days=(rank(date)-1))
ggplot(dd,aes(x=days,y=deaths)) + 
  geom_line(lwd=2,alpha=0.5) + geom_point(size=4) +
  scale_y_continuous(trans='log10',limits = c(10,NA)) +
  geom_abline(aes(intercept=log10(10),slope=log10(2)/2),lty=3) +
  geom_abline(aes(intercept=log10(10),slope=log10(2)/3),lty=2) +
  geom_text(aes(label="2x in 2 days",x=2,y=20,angle=atan2(1,2)/pi*180),hjust=0,vjust=0) +
    geom_text(aes(label="2x in 3 days",x=3,y=20,angle=atan2(1,3)/pi*180),hjust=0,vjust=0) +
  xlab("Days since 10th death") + ylab("Total deaths") +
  theme_bw()
```

Dotted line 2x increase every two days. 

Dashed line 2x increase every three days. 

## South East Michigan Cases
```{r echo=FALSE}
dd <- dc %>% filter(state=="Michigan",county %in% c("Wayne","Oakland","Macomb","Livingston","St. Clair","Lapeer","Genesee","Washtenaw","Monroe","Lenawee")) %>% group_by(date) %>% summarize(deaths=sum(deaths),cases=sum(cases)) %>% filter(cases>=100) %>% mutate(days=(rank(date)-1))

ggplot(dd,aes(x=days,y=cases)) + 
  geom_line(lwd=2,alpha=0.5) + geom_point(size=4) +
  scale_y_continuous(trans='log10',limits = c(100,NA)) +
  geom_abline(aes(intercept=log10(100),slope=log10(2)/2),lty=3) +
  geom_abline(aes(intercept=log10(100),slope=log10(2)/3),lty=2) +
  geom_text(aes(label="2x in 2 days",x=2,y=200,angle=atan2(1,2)/pi*180),hjust=0,vjust=0) +
    geom_text(aes(label="2x in 3 days",x=3,y=200,angle=atan2(1,3)/pi*180),hjust=0,vjust=0) +
  xlab("Days since 100th detected case") + ylab("Detected cases") +
  theme_bw()
knitr::kable(dd)
```

## Deaths in Wayne, Oakland and Macomb
```{r echo=FALSE}
dd <- dc %>% filter(state=="Michigan",county %in% c("Wayne","Oakland","Macomb")) %>% filter(deaths>=10) %>% group_by(county) %>% mutate(days=rank(date)-1)

ggplot(dd,aes(x=days,y=deaths,col=county)) + 
  geom_line(lwd=2,alpha=0.5) + geom_point(size=4) +
  scale_y_continuous(trans='log10',limits = c(10,NA)) +
  geom_abline(aes(intercept=log10(10),slope=log10(2)/2),lty=3) +
  geom_abline(aes(intercept=log10(10),slope=log10(2)/3),lty=2) +
    geom_text(aes(label="2x in 2 days",x=2,y=20,angle=atan2(1,2)/pi*180),hjust=0,vjust=0) +
    geom_text(aes(label="2x in 3 days",x=3,y=20,angle=atan2(1,3)/pi*180),hjust=0,vjust=0) +
  xlab("Days since 10th death") + ylab("Total deaths") +
  theme_bw()
```

## Cases in Wayne, Oakland and Macomb
```{r echo=FALSE}
dd <- dc %>% filter(state=="Michigan",county %in% c("Wayne","Oakland","Macomb")) %>% 
  filter(cases>=100) %>% group_by(county) %>% 
  mutate(days=rank(date)-1)

ggplot(dd,aes(x=days,y=cases,col=county)) + 
  geom_line(lwd=2,alpha=0.5) + geom_point(size=4) +
  scale_y_continuous(trans='log10',limits = c(100,NA)) +
  geom_abline(aes(intercept=log10(min(cases)),slope=log10(2)/2),lty=3) + 
  geom_abline(aes(intercept=log10(100),slope=log10(2)/3),lty=2) + 
    geom_text(aes(label="2x in 2 days",x=2,y=200,angle=atan2(1,2)/pi*180),hjust=0,vjust=0) +
    geom_text(aes(label="2x in 3 days",x=3,y=200,angle=atan2(1,3)/pi*180),hjust=0,vjust=0) +
    xlab("Days since 100th detected case") + ylab("Detected cases") +
  theme_bw()
```
